---
title: "Working with the Madrid Open Data API"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{api-guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE   
)
```

# Overview

This vignette shows how to:

-   Discover API endpoints from the OpenAPI spec
-   Inspect parameters (path vs query)
-   Call endpoints with fetch_api()
-   Turn JSON responses into tables

We’ll use the helper functions included in `datosmadrid`:

-   `list_api_endpoints()` - list and filter endpoints
-   `fetch_api()` - call one endpoint with path_params and query

```{r}
library(datosmadrid)
```

# Discover endpoints

List endpoints and filter by keyword (matches tags, path, summary) and method:

```{r}
# All GET endpoints that mention "bici"
eps <- list_api_endpoints(filter = "deporte", method = "get", include_params = TRUE)
head(eps[, c("path", "method", "title")])
```

# Path vs query parameters

-   **Path params** are placeholders inside the path (e.g., `/{id}` or `.{responseContentType}`), passed via `path_params = list(id = 123, responseContentType = "json")`.

```{=html}
<!-- -->
```
-    **Query params** appear after `?` (e.g., `?q=museo&limite=50`), passed via `query = list(q = "museo", limite = 50)`.

Example templated path: `/catalogo/205099-12104809-aparca-bicis.{responseContentType}`

The `responseContentType` placeholder must be filled with `"json"` (or another allowed value)

# Call API with `fetch_api()`

```{r}
out <- fetch_api(
  path = "/catalogo/210227-0-piscinas-publicas.{formato}",
  path_params = list(formato = "json"),
  parse = TRUE
)
str(out)
```

`parse = TRUE` tells `fetch_api()` to parse JSON and try to return a **tibble** when feasible; otherwise you get a parsed list (for nested structures).

# Find endpoints by category / tag

Filter endpoints by a tag/keyword:

```{r}
cultura <- list_api_endpoints(filter = "cultura", method = "get", include_params = TRUE)
cultura
unique(unlist(cultura$parameters))  # quick glance at parameter names
```
